A Statistical Dynamical Model to Predict Extreme Events and Anomalous Features in Shallow Water Waves with Abrupt Depth Change

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

Controlled laboratory experiments were carried out in  \cite{bolles2018anomalous} to examine the statistical behavior of surface waves crossing an ADC. In these experiments, nearly unidirectional waves are generated by a paddle wheel and propagate through a long, narrow wave tank. Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition to a shallower depth. The paddle wheel is forced with a pseudo-random signal intended to mimic a Gaussian random sea. In particular, the paddle angle is specified as
%The angle of the paddle wheel $\theta(t)$ is prescribed as a pseudo-random signal that mimics a Gaussian random sea 
\[
\theta\left(t\right)=\theta_{0}+\Delta\theta\sum_{n=1}^{N}a_{n}\cos\left(\omega_{n}t+\delta_{n}\right),\;E\sim\left(\Delta\theta\right)^{2}\sum_{n}a_{n}^{2}\omega_{n}^{2}.
\]
where the weights $a_n$ are Gaussian in spectral space. The peak frequency $\omega_p$ gives rise to a characteristic wavelength $\lambda_c$ of water waves, which can be estimated using the dispersion relation.
Above, the total energy $E$ in the system is determined by the angle amplitude
$\Delta\theta$. Optical measurements of the free surface reveal a number of surprising statistical features:
\begin{itemize}
\item Distinct statistics are found between the incoming and outgoing wave
disturbances: incoming waves display near-Gaussian statistics,
while outgoing waves skew strongly towards positive displacement.
\item The degree of non-Gaussianity present in the outgoing wave field depends on the total energy of the system: 
larger driving amplitudes $\Delta\theta$ generate stronger skewness in the surface displacement PDFs.
\item \textcolor{blue}{The waves also show different slopes in the power spectra at upstream and downstream locations}.

\todo[inline]{I am not sure we need to include this last item. It is not really an experimental observation, simply a calculation that can be done using the dispersion relation and the values of the two depths. Could we simply take this last item out? Would you prefer to replace it with another observation? I could easily come up with one. Perhaps a comment on the decay rate of the power spectra. (I will confirm with Andy whether he would just delete this item or keep something else like the one above)}

>>>>

Controlled laboratory experiments were carried out in  \cite{bolles2018anomalous} to examine the statistical behavior of surface waves crossing an ADC. In these experiments, nearly unidirectional waves are generated by a paddle wheel and propagate through a long, narrow wave tank. Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition to a shallower depth. The paddle wheel is forced with a pseudo-random signal intended to mimic a Gaussian random sea upstream of the ADC. In particular, the paddle angle is specified as
\[
\theta\left(t\right)=\theta_{0}+\Delta\theta\sum_{n=1}^{N}a_{n}\cos\left(\omega_{n}t+\delta_{n}\right),\;E\sim\left(\Delta\theta\right)^{2}\sum_{n}a_{n}^{2}\omega_{n}^{2}.
\]
where the weights $a_n$ are Gaussian in spectral space with peak frequency $\omega_p$ and the phases $\delta_n$ are uniformly distributed random variables. The peak frequency gives rise to a characteristic wavelength $\lambda_c$ which can be estimated from the dispersion relation. The energy $E$ injected into the system is determined by the angle amplitude $\Delta\theta$, which is the main control parameter varied in \cite{bolles2018anomalous}. Optical measurements of the free surface reveal a number of surprising statistical features:
\begin{itemize}
\item Distinct statistics are found between the incoming and outgoing wave
disturbances: incoming waves display near-Gaussian statistics,
while outgoing waves skew strongly towards positive displacement.
\item The degree of non-Gaussianity present in the outgoing waves depends on the injected energy $E$: 
larger energies generate stronger skewness in the surface displacement PDFs and more extreme events.
\item \textcolor{blue} {Compared to the incoming wave train, the power spectrum of the outgoing wave field decays more slowly, which indicates that the anomalous behavior is associated with an elevated level of high frequencies.}
%{The waves also show different slopes in the power spectra at upstream and downstream locations}.
\end{itemize}
